---
title: "Graphab R Java Integration"
author: "Sebastián Sánchez Cuartas"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    theme: united
    highlight: tango
    css: styles.css
---

# **Introduction**

The following code integrates Java within R using the Graphab software. In this case, the **´rJava´** package is used to execute the Graphab **´.jar´** file, which must be downloaded from its [official website](https://sourcesup.renater.fr/www/graphab/en/home.html). In this example, version 2.8.6 was used. Additionally, a cost surface or cost raster is utilized, along with a raster containing the patches within the landscape, similar to how it would be done with the Graphab software. A specific configuration is employed, but it can be modified according to the needs of the case. All related information can be found in the Graphab manual. The code is also structured within a **´for´** loop, allowing the process to be automated for different cost surfaces. In this case, different cost surfaces are used with the same patches, although it is possible to include different patches for each resistance surface if necessary.

```{r eval=FALSE}

library(rJava)

# Define the path to the JAR file
jar_path <- "C:/Users/Usuario/Desktop/Archivos_tesis/graphab-2.8.6.jar"

# Load the JAR file
.jinit()
.jaddClassPath(jar_path)

# Specify the path to the folder containing the .tif files
files_folder <- "C:/Users/Usuario/Desktop/Archivos_tesis/RC_WGS_32618/RC_WGS_32618_with_buffers/" # Contains the rasters with patches
cost_folder <- "C:/Users/Usuario/Desktop/Archivos_tesis/RC_WGS_32618/" # Contains the cost rasters

# Specify the path to the folder where Graphab results will be saved
results_folder <- "C:/Users/Usuario/Desktop/Archivos_tesis/Graphab_resultados/"

# Get the list of .tif files in the files folder
tif_files <- list.files(files_folder, pattern = "\\.tif$", full.names = TRUE)
cost_files <- list.files(cost_folder, pattern = "\\.tif$", full.names = TRUE)

# Sort the files alphabetically to ensure they are paired correctly, so that the for loop matches the cost raster with the corresponding patch raster. It is important that the filenames have some distinctive characteristic that allows proper sorting.

tif_files <- sort(tif_files)
cost_files <- sort(cost_files)

# Check that the number of files matches
if (length(tif_files) != length(cost_files)) {
  stop("The number of input files does not match.")
}

# Iterate over the files in pairs
for (i in seq_along(tif_files)) {
  # Get the current file and its corresponding cost file
  tif_file <- tif_files[i]
  cost_file <- cost_files[i]
  
  # Get the iteration number for the Conectparamo folder name
  iteration_number <- i
  
  # Get the filename without the path
  file_name <- basename(cost_file)
  
  # Remove the file extension
  name_without_extension <- tools::file_path_sans_ext(file_name)
  
  # Create the corresponding Conectparamo folder name
  conectparamo_folder <- file.path(results_folder, name_without_extension)
  
  # Create the Conectparamo folder if it does not exist
  if (!dir.exists(conectparamo_folder)) {
    dir.create(conectparamo_folder, recursive = TRUE)
  }
  
  # Build the arguments to call the main method of the JAR file
  params <- c(
    "--create", "Conectparamo",
    tif_file,
    "habitat=150,153",
    "nodata=255",
    "minarea=1",
    "con8",
    paste0("dir=", conectparamo_folder),
    "--linkset", "distance=cost", "name=linkset1", "complete", "maxcost=100000000", 
    paste0("extcost=", cost_file),
    "--graph", "name=graph1", "threshold=100000000",
    "--usegraph", "graph1",
    "--lmetric", "CF", "beta=0.0",
    "--lmetric", "CF", "beta=1.0"
  )
  
  # Execute the main method of the JAR file with the defined parameters
  .jcall("org.thema.graphab.MainFrame", "V", "main", params)
}

```

## **Linksets in New Folder**

After generating the models using Graphab, the results are organized into separate folders for each resistance surface. Within these folders, there is a .csv file called **´linkset´**, which contains the calculated connectivity indices for each pair of patches studied. The following code takes all the linkset files and renames them, assigning each a unique name that references the cost raster used. Additionally, a numeric prefix (optional) is added, ranging from 1 to the total number of files. This numeric prefix is useful for the historical connectivity algorithm to correctly identify the connectivity indices of each cost raster and perform the corresponding calculations ([see the algorithm here](https://github.com/innerhaze/R-algorithm-for-Historical-Connectivity-Analysis-for-P-ramo-Lineages-using-a-Pleistocene-polen-model)). It is important to set the path of the linksets within the **´archivos_indices´** vector in the algorithm's code so that they can be used correctly.

```{r eval=FALSE}

# Define the path to the main results folder
main_path <- "C:/Users/Usuario/Desktop/Archivos_tesis/Graphab_resultados"

# Path to the results folder for the ordered files
output_path <- file.path(main_path, "Ordered_lmetrics_Results")

# Create the results folder if it does not exist
if (!dir.exists(output_path)) {
  dir.create(output_path)
}

# Get the list of subfolders within the main folder
subfolders <- list.dirs(path = main_path, full.names = TRUE, recursive = FALSE)

# Initialize counter for enumeration (numeric prefix)
counter <- 1

# Iterate over the subfolders
for (subfolder in subfolders) {
  # Get the name of the subfolder
  subfolder_name <- basename(subfolder)
  
  # Get the path to the "Conectparamo" folder within the subfolder
  conectparamo_folder <- file.path(subfolder, "Conectparamo")
  
  # Get the list of CSV files within the "Conectparamo" folder
  csv_files <- list.files(path = conectparamo_folder, pattern = "linkset1-links.csv", full.names = TRUE)
  
  # Iterate over the CSV files and process them
  for (csv_file in csv_files) {
    # Read the CSV file
    data <- read.csv(csv_file, stringsAsFactors = FALSE)
    
    # Construct the output file name with enumeration at the beginning
    output_file_name <- paste0(counter, "_", subfolder_name, "_", basename(csv_file))
    
    # Write the (ordered) data to a new CSV file in the results folder
    write.csv(data, file = file.path(output_path, output_file_name), row.names = FALSE)
    
    # Increment counter
    counter <- counter + 1
  }
}

```
